\textbf{\sc Deliberative Autonomy for Marine Robotics}

### 2012-01-30 trunk/europa.tex
checkpoint. more europa content.

Scheduling:
In the Resource Constrained Project Scheduling Problem (RCPSP), a project consisting of a set of activities must be scheduled in a way that satisfies minimum and/or maximum temporal separation constraints among them, as well as respects fixed limits on the availability of resources required to perform each activity. In addition to satisfying temporal and resource constraints, it is common for the user to want to minimize makespan so that the entire project is finished as early as possible.

@@

Planning:
In the Shopping Agent Problem (TODO: ref to Norvig) an agent needs to purchase a set of products that are available at specific locations (supermarket, hardware store, etc), the agent  is subject to temporal (must complete tasks by specific deadlines) and resource (fuel, carrying capacity, etc) constraints.

@@

As we can see, a planning problem (what actions must be performed to achieve a goal) may embed a scheduling problem (in what order should those actions be performed), and both embed Constraint Satisfaction problems since temporal, resource and many other kinds of constraints must be respected in most real life problems.

@@

CP is normally implemented as part of a programming language and constraints are normally represented as objects [PUG95]. Any constraint can be introduced and, as long as the bounds propagation and domain reduction protocols specified by the host CP system are enforced, it will be indistinguishable from any other “primitive” CP constraint, such as <= or >=.

@@

Shopping model:

@@

And a problem instance:

@@

// Indicate that the database is closed - no new objects can be created
close();

@@

TODO: introduce model for Shopping link to architectural components.
explain variable, object, predicate and action types and instances, constraints, goals and facts.

@@

TODO: explain how inference is used to detect perform bounds propagation, domain reduction and to detect constraint violations.
- talk about AC3 implementation and extension mechanisms
- talk about europa's constraint library
- talk about Temporal Network
- talk about Resource Envelopes.

@@

	\item Open Condition: an open condition is an inactive token. Inactive tokens can be generated by explicitly posted goals, or when a token is activated, rules may fire that create inactive slave tokens (TODO: point to more detailed decription above in Rules Engine section).  Open Conditions can be resolved in three ways: 
\begin{enumerate}
		\item Merging: A free activity is merged with a matching activity already in the plan. An activity a is said to match an activity a? if a and a? unify and the temporal constraints involving a are satisﬁed by a?. Thus,  a and a? can be considered the same activity and we do not need to introduce a in the plan. Consequently, the compatibilities associated with a are not ﬁred, because they have been already triggered when a? was introduced in the plan. 
		\item Activation: We introduce a new activity a in the current plan associating it with the proper timeline, but without choosing a speciﬁc time slot for it.The compatibilities associated with a are applied and the subgoal activities resulting from those compatibilities are introduced as free activities. This results in both an ordering ﬂaw, corresponding to the just activated activity, and a number of open condition ﬂaws, corresponding to the new subgoal activities. 
		\item Rejection: Some goals may be optional, if that is the case for a particular order condition, its corresponding flaw can be resolved by discarding the inactive token.
\end{enumerate}

@@

>>>>

Scheduling:
In the Resource Constrained Project Scheduling Problem (RCPSP), a project consisting of a set of activities must be scheduled in a way that satisfies minimum and/or maximum temporal separation constraints. The activity schedule must also respect fixed limits on the availability of resources required to perform each activity. In addition to satisfying temporal and resource constraints, it is common for the user to want to minimize makespan so that the entire project is finished as early as possible.

@@

Planning:
In the Shopping Agent Problem (TODO: ref to Norvig) an agent needs to purchase a set of products that are available at specific locations (supermarket, hardware store, etc), the agent  is subject to temporal (must complete tasks by specific deadlines) and resource (fuel, carrying capacity, etc) constraints. The agent must figure out what actions need to be performed to find and acquire the required items, as well as when to perform each of those actions.

@@

As we can see, a planning problem (what actions must be performed to achieve a goal) may embed a scheduling problem (in what order should those actions be performed), and both planning and scheduling embed Constraint Satisfaction problems since temporal, resource and many other kinds of constraints must be respected in most real life problems.

@@

CP is normally implemented as part of a programming language and constraints are normally represented as objects [PUG95]. Any constraint can be introduced and, as long as the bounds propagation and domain reduction protocols specified by the host CP system are enforced, it will be indistinguishable from any other ÒprimitiveÓ CP constraint, such as <= or >=.

Solving a CSP is NP-Hard in theory, but often very efficient in practice using a number of algorithms and techniques that are well understood.

@@

EUROPA can ingest descriptions of models, plans and goals. This means that a single planner can be applied to different systems and problems, simply by providing different models and goals. For such planners the expressiveness of the languages for models, goals and plans is of great importance. EUROPA provides a high level language called NDDL (New Domain Description Language) that allows the user to specify the different components of a problem domain in a concise way. The main characteristics that make NDDL a powerful tool for describing problem domains are:

\begin{enumerate}
	\item It's Object Oriented: NDDL supports classes and inheritance in similar fashion to popular Object-Oriented languages like Java and C++. It also supports polymorphism for some of the planning components. Using object classes and instances is a time-tested approach to naturally describe problem domains (TODO: OO ref?).
	\item It offers declarative constructs to define constraints and causality dependencies between actions and effects.
	\item It offers procedural constructs to populate EUROPA's plan database with specific problem instances expressed in terms of objects, variables, constraints, facts and goals.
\end{enumerate}

To understand how the most important elements of NDDL work, let's take a look at how the Shopping Agent problem described above would be modeled:

@@

TODO: explain variable, object, predicate and action types, including constraints that can be part of them

In this model we can see:

\begin{enumerate}
	\item Variable Types:
	\item Object Types:
	\item Predicate Types:
	\item Action Types:
\end{enumerate}

Now let's look at how a specific problem instance would be specified in NDDL:

@@

// Indicate that the database is closed - no new objects can be created
// this allows EUROPA to perform more efficient inference
close();

@@

TODO:  explain variable, object, predicate and constraint instances.
explain goals and facts.

\begin{enumerate}
	\item Variable Instances:
	\item Object Instances:
	\item Facts:
	\item Goals:
	\item Constraint Instances:
\end{enumerate}

@@

TODO: explain how inference is used to detect perform bounds propagation, domain reduction and to detect constraint violations.
- talk about AC3 implementation and extension mechanisms
- talk about europa's constraint library

The Temporal Network:
The notion of time is central to temporal planning. EUROPA uses variables to explicitly represent timepoints for plan activities and states. Constraints among timepoints provide a natural way to express domain axioms. For example, in order to state that activity A must occur before activity B we can say that the end timepoint of A is <= the start timepoint of B. Dechter et al. (Dechter 1991) proposed that constraints among timepoints can be grouped together to form a Simple Temporal Network (STN). Such a network can be transformed into a Distance Graph (DG) where the outward arc from a node represents the maximum distance from the source node to the target node. The diagram below illustrates a simple STN with just 2 variables and a single constraint. It also shows the resulting DG.

TODO: add figure

Dechter et al. (Dechter 1991) also showed that shortest path algorithms could be used to propagate values in the network and discover a negative cycle. A negative cycle is a path from a node to itself that has a path length less than 0. If such a cycle exists, the network is inconsistent. It was further shown that a single-source shortest path algorithm was sufficient to detect a negative cycle and provide sufficient propagation to yield a backtrack-free search. Thus we have an efficient and complete algorithm for propagating an STN. These results build on the already established notion of a CSP and are naturally incorporated into the general representation and propagation scheme used in EUROPA.

Resources:

    Reservoir resources can have both 'consume' and 'produce' tokens. Note that these tokens do NOT have the usual start/end/duration token variables, since occur at a single instant in time. Instead, they have a single variable, 'time' used to represent that instant.
    Reusable resources have 'uses' tokens that use a quantity for the duration of the token. These tokens do have the usual start/end/duration variables.
    Unary resources are reusable resources with unit quantity and have 'use' tokens (again, with the usual start/end/duration variables).

    TODO: talk about resource envelope computation by using MaxFlow formulation.

@@

	\item Open Condition: an open condition is an inactive token. Inactive tokens can be generated by explicitly posted goals, or when a token is activated, rules may fire that create inactive slave tokens (TODO: point to more detailed decription above in Rules Engine section).  Open Conditions can be resolved in three ways: 
\begin{enumerate}
		\item Merging: A free activity is merged with a matching activity already in the plan. An activity a is said to match an activity a? if a and a? unify and the temporal constraints involving a are satisÞed by a?. Thus,  a and a? can be considered the same activity and we do not need to introduce a in the plan. Consequently, the compatibilities associated with a are not Þred, because they have been already triggered when a? was introduced in the plan. 
		\item Activation: We introduce a new activity a in the current plan associating it with the proper timeline, but without choosing a speciÞc time slot for it.The compatibilities associated with a are applied and the subgoal activities resulting from those compatibilities are introduced as free activities. This results in both an ordering ßaw, corresponding to the just activated activity, and a number of open condition ßaws, corresponding to the new subgoal activities. 
		\item Rejection: Some goals may be optional, if that is the case for a particular order condition, its corresponding flaw can be resolved by discarding the inactive token.
\end{enumerate}

@@

N-Queens solver:
\begin{verbatim}
    public void solve(int maxIter)
    {
    	init();

    	for (int i=0;psengine_.getViolation() > 0 && i < maxIter;i++) {
    	    PSVariable queenToMove = getQueenWithMaxViolation();
    		int curPos = queenToMove.getSingletonValue().asInt();

    	    boolean moved = false;
            SortedSet<Move> moves = getMoves(queenToMove,curPos);             
            for (Move m : moves) {
            	moved = makeMove(queenToMove,curPos,m,false);
            	if (moved)
            	    break;
            }

            if (!moved) 
            	makeMove(queenToMove,curPos,moves.first(),true);

            checkSolution(); // See if we have a new best solution            
            notifyIterationCompleted(curIteration_++);

            if (curIteration_-bestIter_ > 50) 
            	restart();
    	}
    }
\end{verbatim}

RCPSP solver:
\begin{verbatim}
    public void solve(PSEngine psengine,
    		          long timeout, // in msecs
    		          int bound)
    {
       init(psengine,timeout,bound);

       for (int i=0; true ; i++) {
          flatten();
          updateSolution(i);
          updateCriticalPrecedences();

          if ((nbStable_ > maxStable_) || (bestMakespan_ <= makespanBound_))
             break;

          if (timer_.getElapsed() > timeout) {
              timedOut_ = true;
              break;
          }

          relax();
          curIteration_++;
       }

       restoreBestSolution();
    }
